---
title: "COVID-19 Growth"
---

## Submitted Date: 04/01/2020

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r warning=FALSE}
suppressPackageStartupMessages(library(rgdal))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(leaflet))
```

```{r cache=TRUE}
# Boundary data from https://github.com/datasets/geo-countries
# downloaded 03/31/2020

countries <- readOGR("../../data/countries.geojson", verbose = FALSE)
codata <- countries@data
```

```{r cache=TRUE}

# Downloaded COVID data 04/01/20, 8:30 AM EDT.

# download.file("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv", "data/casedistribution.csv", mode = "wb")

cov <- read.csv("data/casedistribution.csv", stringsAsFactors = FALSE)
# Fixes for known differences in codes
cov$geoId[cov$geoId == "UK"] <- "GB"
cov$geoId[cov$geoId == "EL"] <- "GR"
names(cov)[1] <- "date"
cov$date <- dmy(cov$date)
```

```{r}
codata$ISO_A2 <- as.character(codata$ISO_A2)
last.report <- max(cov$date)
base.end <- last.report - 7
base.start <- base.end - 13
covuniq <- sort(unique(cov$geoId))
cov.base <- sapply(covuniq, function(x) sum(filter(cov, between(date, base.start, base.end), geoId == x)$cases))
cov.add <- sapply(covuniq, function(x) sum(filter(cov, date > base.end, geoId == x)$cases))
cov.growth <- cov.add / (cov.base)
cov.data <- data.frame(
    ISO_A2 = covuniq,
    base = as.vector(cov.base),
    add = as.vector(cov.add),
    growth = as.vector(cov.growth),
    stringsAsFactors = FALSE
)
suppressMessages(codata <- codata %>% left_join(cov.data))
codata$base[is.na(codata$base)] <- 0
codata$add[is.na(codata$add)] <- 0
codata$growth[is.na(codata$growth) | codata$growth == Inf] <- 0
```

```{r}
bins <- unique(quantile(codata$growth, seq(0,1, by = 0.1)))
palbin <- colorBin(
    palette = "YlOrRd",
    bins <- length(bins),
    domain = bins
)
```

## COVID-19 Growth Rate

```{r}
labels <- sprintf("<strong>%s</strong><br />New Case Rate: %g%%<br />Prev: %g / New: %g", codata$ADMIN, round(codata$growth * 100, 1), codata$base, codata$add) %>% lapply(htmltools::HTML)
m4 <- leaflet(countries) %>% addPolygons(weight = 1, color = "black", fillColor = ~palbin(codata$growth), fillOpacity = 1, label = labels)
m4
```

Growth is the ratio of cases reported in the last week divided by the total reported cases in the previous two weeks.  A steady state would have a rate of 50% by this calculation.  The two week window is used as a proxy for the number of active cases at the start of the last week.

Map data is from Boundary data from Natural Earth via https://github.com/datasets/geo-countries.

COVID-19 data is from European Centre for Disease Prevention and Control, https://www.ecdc.europa.eu/en/publications-data/download-todays-data-geographic-distribution-covid-19-cases-worldwide.

NOTE: This is done with Leaflet, but does not use the standard Leaflet tileset.

